---
title: "DATA605 WK10 Assignment 10"
author: "Javern Wilson"
date: "10/28/2019"
header-includes:
  - \usepackage{blkarray}
  - \usepackage{amsmath}
output:
    html_document:
      theme: readable
---


## Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars. A guard agrees to make a series of bets with him. If Smith bets A dollars, he wins A dollars with probability `.4` and loses A dollars with probability `.6`. 
 
Find the probability that he wins 8 dollars before losing all of his money if 
 
 
### (a) he bets 1 dollar each time (timid strategy). 

Smith's outcome is a Markov Chain

Let's create the transition matrix in `canonical form`.


P =
\begin{array}{c | c}
  Q & R \\
  \hline
  0 & I\\
\end{array}\\


P =
\begin{array}{c c c c c c c c | c c }
  & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 0 & 8 \\
  1 & 0 & 0.4 & 0 & 0 & 0 & 0 & 0 & 0.6 & 0\\
  2 & 0.6 & 0 & 0.4 & 0 & 0 & 0 & 0 & 0 & 0 \\
  3 & 0 & 0.6 & 0 & 0.4 & 0 & 0 & 0 & 0 & 0\\
  4 & 0 & 0 & 0.6 & 0 & 0.4 & 0 & 0 & 0 & 0\\
  5 & 0 & 0 & 0 & 0.6 & 0 & 0.4 & 0 & 0 & 0\\
  6 & 0 & 0 & 0 & 0 & 0.6 & 0 & 0.4 & 0 & 0\\
  7 & 0 & 0 & 0 & 0 & 0 & 0.6 & 0 & 0 & 0.4\\
  \hline
  0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0\\
  8 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 \\
\end{array}

From this we see that the matrix **Q** is

Q = \begin{pmatrix} 
0 & 0.4 & 0 & 0 & 0 & 0 & 0\\
0.6 & 0 & 0.4 & 0 & 0 & 0 & 0 \\
0 & 0.6 & 0 & 0.4 & 0 & 0 & 0\\
0 & 0 & 0.6 & 0 & 0.4 & 0 & 0\\
0 & 0 & 0 & 0.6 & 0 & 0.4 & 0\\
0 & 0 & 0 & 0 & 0.6 & 0 & 0.4\\
0 & 0 & 0 & 0 & 0 & 0.6 & 0
\end{pmatrix}

I = \begin{pmatrix} 
1 & 0 & 0 & 0 & 0 & 0 & 0\\
0 & 1 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 1 & 0 & 0 & 0 & 0\\
0 & 0 & 0 & 1 & 0 & 0 & 0\\
0 & 0 & 0 & 0 & 1 & 0 & 0\\
0 & 0 & 0 & 0 & 0 & 1 & 0\\
0 & 0 & 0 & 0 & 0 & 0 & 1
\end{pmatrix}

R = \begin{array}{ccc}
& 0 & 8 \\
1 & 0.6 & 0 \\
2 & 0 & 0 \\
3 & 0 & 0\\
4 & 0 & 0\\
5 & 0 & 0\\
6 & 0 & 0\\
7 & 0 & 0.4\\
\end{array}


Calculation:

```{r}
I <- diag(7)
I


Q <- matrix(c(rep(c(0, 0.4, 0, 0, 0, 0, 0, 0.6), 6), 0), 7, byrow = T)

Q

I - Q

# Fundamental Matrix
N <- solve(I - Q)
round(N, 4)

R <- matrix(rep(c(0, 0), 7), nrow = 7)
R[1,1] <- 0.6
R[7,2] <- 0.4
R

B <- N %*% R
colnames(B) <- c(0, 8)
round(B, 4)

```


### (b) he bets, each time, as much as possible but not more than necessary to bring his fortune up to 8 dollars (bold strategy). 

In other words with each bet, his winning doubles.
The chances of this happening is like so:

Smith starts with 1

Risk 1: He wins 2

Risk 2: He wins 4

Risk 3: He wins 8

Just 3 chances (in a row) he may need to take using the bold strategy. The chances of luck being on Smith's side is:

```{r}
0.4^3

```


### (c) Which strategy gives Smith the better chance of getting out of jail? 
The bold strategy with `6.4%` compared to the timid strategy with `2.03%`