---
title: "DATA605 WK4 Assignment4"
author: "Javern Wilson"
date: "9/10/2019"
output:
  html_document: default
  pdf_document: default
---


```{r message=TRUE, warning=TRUE}
library(Matrix)
library(matrixcalc)
library(matlib)
```


### Problem Set 1

1. What is the rank of the matrix A? 

$$\mathbf{A} =\left[\begin{array}
{rrr}
1 & 2 & 3 & 4\\
-1 & 0 & 1 & 3\\
0 & 1 & -2 & 1\\
5 & 4 & -2 & -3
\end{array}\right]$$

```{r}

A <- matrix(c(1, 2, 3, 4, -1, 0, 1, 3, 0, 1, -2, 1, 5, 4, -2, -3), 4, 4, byrow = T)

```

Using Row Echelon form
```{r}
#echelon
echelon(A, reduced = F)
```
Counting all non-zero rows tells the rank of the matrix. In this case it is `4`.

To confirm this using R's built in function:
```{r}
rankMatrix(A)
```
The rank is `4`.


2. Given an mxn matrix where **m > n**, what can be the maximum rank? The minimum rank, assuming that the matrix is non-zero? 

    - If m is less than n, then the maximum rank of the matrix is m.
    - If m is greater than n, then the maximum rank of the matrix is n.
    - The minimum rank would be one.
    

3. What is the rank of matrix B?

  
$$\mathbf{B} =\left[\begin{array}
{rrr}
1 & 2 & 1\\
3 & 6 & 3\\
2 & 4 & 2\\
\end{array}\right]$$

```{r}
B <- matrix(c(1, 2, 1, 3, 6, 3, 2, 4, 2), 3, 3, byrow = T)

```


```{r}
echelon(B, reduced = F)

```
The rank is `1`.



### Problem Set 2

1. Compute the eigenvalues and eigenvectors of the matrix A. You’ll need to show your work. You’ll need to write out the characteristic polynomial and show your solution.

  
$$\mathbf{A} =\left[\begin{array}
{rrr}
1 & 2 & 3\\
0 & 4 & 5\\
0 & 0 & 6\\
\end{array}\right]$$

```{r}
A <- matrix(c(1, 2, 3, 0, 4, 5, 0, 0, 6), 3, 3, byrow = T)
I <- diag(3)
```

To obtain the eigenvalues of A, we find the determinant >> $det(A - \lambda I) = 0$ to satisfy the characteristic equation of the matrix A.

  
\[\begin{aligned} \mathbf{A - \lambda I} &=\left[\begin{array}
{rrr}
1 & 2 & 3\\
0 & 4 & 5\\
0 & 0 & 6\\
\end{array}\right] -   
\left[\begin{array}
{rrr}
\lambda & 0 & 0\\
0 & \lambda & 0\\
0 & 0 & \lambda\\
\end{array}\right]  
=\left[\begin{array}
{rrr}
1 - \lambda & 2 & 3\\
0 & 4 - \lambda & 5\\
0 & 0 & 6 - \lambda\\
\end{array}\right]\\ \\ \\

\mathbf{det(A - \lambda I)} &= (1- \lambda) \begin{vmatrix}
4 - \lambda & 5\\
0 & 6 - \lambda
\end{vmatrix} - (2)\begin{vmatrix}
0 & 5\\
0 & 6 - \lambda
\end{vmatrix} +(3)\begin{vmatrix} 0 & 4 - \lambda\\
0 & 0
\end{vmatrix}\\ \\
&= (1 - \lambda)\begin{vmatrix}(4 - \lambda)(6-\lambda) - (0)(5)\end{vmatrix} - (2) \begin{vmatrix}(0)(6- \lambda) - (0)(5)\end{vmatrix} + (3) \begin{vmatrix}(0)(0)- (0)(4- \lambda)\end{vmatrix}\\ \\
&= (1 - \lambda)(4 - \lambda)(6-\lambda) - 0 + 0\\ \\
&= (1 - \lambda)(4 - \lambda)(6-\lambda)\end{aligned}\]

The characteristic polynomial is $-\lambda^3 + 11 \lambda^2 - 34\lambda + 24$

Therefore the eigenvalues of $A$ are $\lambda$ = `1`, `4` and `6`


Eigenvectors:

#### $\lambda = 1$

$-A + I_3$


```{r}
-A + I

```

```{r}
echelon((-A + I))
```

$$\left[\begin{array}
{rrr}
1\\0\\0\\
\end{array}\right]$$


#### $\lambda = 4$

$-A + 4I_3$

```{r}
-A + (4*I)
```


```{r}
echelon((-A + (4*I)))
```


$$\left[\begin{array}
{rrr}
\frac{2}3\\1\\0\\
\end{array}\right] = \left[\begin{array}
{rrr}
2\\3\\0\\
\end{array}\right]$$


#### $\lambda = 1$

$-A + 6I_3$

```{r}

-A + (6*I)

```

```{r}
echelon((-A + (6*I)))
```


$$\left[\begin{array}
{rrr}
\frac{8}5\\\frac{5}2\\1\\
\end{array}\right] = \left[\begin{array}
{rrr}
16\\25\\10\\
\end{array}\right]$$


#### Helpful Links:

[FINDING EIGENVALUES AND EIGENVECTORS](https://www.scss.tcd.ie/~dahyotr/CS1BA1/SolutionEigen.pdf)

[Eigenvalues and Eigenvectors](https://lpsa.swarthmore.edu/MtrxVibe/EigMat/MatrixEigen.html)