---
title: "DATA605 Wk2Assignment2"
author: "Javern Wilson"
date: "9/7/2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
---


### (1) Show that $A^TA \neq AA^T$ in general. (Proof and demonstration.) 

Let Matrix A be $$ \mathbf{A} = \left[\begin{array}
{rrr}
a & r\\
b & s
\end{array}\right]$$ 

and

$$\mathbf{A^T} = \left[\begin{array} {rrr} a & b\\ r & s\end{array}\right]$$ 
Proof:

$$\mathbf{A^TA} = \left[\begin{array}
{rrr}
a \times a + b \times b & a \times r + b \times s\\ 
r \times a + s \times b & r \times r + s \times s
\end{array}\right]$$

$$\mathbf{A^TA} = \left[\begin{array}
{rrr}
a^2 + b^2 & ar + bs\\ 
ra + sb & r^2 + s^2
\end{array}\right]$$

$$\mathbf{AA^T} = \left[\begin{array}
{rrr}
a \times a + r \times r & a \times b + r \times s\\ 
b \times a + s \times r & b \times b + s \times s
\end{array}\right]$$

$$\mathbf{AA^T} = \left[\begin{array}
{rrr}
a^2 + r^2 & ab + rs\\ 
ab + sr & b^2 + s^2
\end{array}\right]$$


Therefore $A^TA \neq AA^T$.

For example:

*A*
```{r}

A = matrix(c(1, -2, 4, -3), 2, 2, byrow = T)
A


```

*A Transpose*
```{r}
TA = t(A)
TA

```

*A multiplied by A Transpose*
```{r}

A %*% TA
```


*A Transpose multiplied by A*
```{r}
TA %*% A
```



### (2) For a special type of square matrix A, we get $A^TA = AA^T$ . Under what conditions could this be true? (Hint: The Identity matrix I is an example of such a matrix). 

Conditions:

A matrix is considered symmetric if it is equal to it's transpose.
If $A$ is symmetric then $A = A^T$ so $A^TA$ should be the same as $AA^T$

Proof using rhs

$$AA^T = (AA^T)^T\\$$

$$ = A^T(A^T){^T}\\$$

$$ = A^TA$$

---

A
```{r}

A <- matrix(c(3, 2, 4, 2, 6, 2, 4, 2 ,3), 3, 3, byrow = T)
A


```

*A Transpose*
```{r}

TA <- t(A)
TA


```

*A multiplied by A Transpose* -> $AA^T$
```{r}

A %*% TA

```

*A Transpose multiplied by A* -> $A^TA$

```{r}

TA %*% A

```

Transpose of $AA^T$ rhs
```{r}
t(A %*% TA)

```



### Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer. 

```{r}
LU_D <- function(M){

  m <- dim(M)
  if (m[1] != m[2]){
    print("This is not a square matrix!")
  }
  
  U <- matrix( 0, nrow = nrow(M), ncol=nrow(M) )
  L <- diag(3)
  for(i in 1 : nrow(M)) {
    p1 <- i+1
    m1 <- i-1
    for (j in 1:nrow(M)){
      U[i, j] <- M[i,j]
        if ( m1 > 0 ) {
          for ( k in 1:m1 ) {
            U[i,j] <- U[i,j] - L[i,k] * U[k,j]
          }
        }
      }
    
  if ( p1 <= nrow(M) ) {
    for ( j in p1:nrow(M) ) {
      L[j,i] <- M[j,i]
      if ( m1 > 0 ) {
        for ( k in 1:m1 ) {
          L[j,i] <- L[j,i] - L[j,k] * U[k,i]
        }
      }
      L[j,i] <- L[j,i] / U[i,i]
    }
  }
  }
  
  output <- list(L=L, U=U)
  return(output)
}

```


Test function
```{r}

D <- matrix(c(1, 4, -3, -2, 8, 5, 3, 4, 7), 3, 3, byrow = T)
E <- matrix(c(2, 4, -4, 1, -4, 3, -6, -9, 5), 3, 3, byrow = T)

LU_D(D)
LU_D(E)
```


Helpful Links:

  +[LU Decomposition - Shortcut Method](https://www.youtube.com/watch?v=UlWcofkUDDU)
  
  +[Trackable R function for LU factorization
](https://stackoverflow.com/questions/51687808/write-a-trackable-r-function-that-mimics-lapacks-dgetrf-for-lu-factorization)

  +[The LU Decomposition of a Matrix Examples 1](http://mathonline.wikidot.com/the-lu-decomposition-of-a-matrix-examples-1)
