---
title: "DATA605 WK7 Assignment 7"
author: "Javern Wilson"
date: "10/12/2019"
output: 
  html_document:
    theme: flatly
    highlight: kate
---

## 1.  Let $X_1$, $X_2$,..., $X_n$ be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the $X_i$’s. Find the distribution of Y. 

Solution: For $1 ≤ j ≤ k$, $m(j) = \frac{(k−j+1)^n−(k−j)}{k^n}$

**In other words**

As Y is the minimum value of $X_i$'s over all of the $X_i$'s then the distribution function would be $m(j) = P(Y = j)$.

We will need to count the number of ways that we can assign the $X_i$'s to values between $j$ and $k$ with at least one $X_i$ being assigned to $j$ divided by the total number of possible ways to assign $X_i$'s to values between $1$ and $k$.

 
## 2.  Your organization owns a copier (future lawyers, etc.) or MRI (future doctors).  This machine has a manufacturer’s expected lifetime of 10 years.  This means that we expect one failure every ten years.  (Include the probability statements and R Code for each part.). 


 
### a.  What is the probability that the machine will fail after 8 years?.  Provide also the expected value and standard deviation.  Model as a geometric.  (Hint: the probability is equivalent to not failing during the first 8 years..) 
 
 $P(X=n) = q^{n-1}p$ where $n = 8$ and $p = 0.1$
```{r}
# The probabilty of failure = 1/10
p = 0.1

# n trials
n = 8


dgeom(x = n, prob = p)

```
 
 $E(X) = \frac{1}{p}$
 
```{r}
1 / p
```
 
$D(X) = \frac{1-p}{p^2}$
```{r}
sqrt((1 - p) / p**2)

```

 
### b.  What is the probability that the machine will fail after 8 years?.  Provide also the expected value and standard deviation.  Model as an exponential.
  
  $P(X \ge n) = e^{\frac{-k}{\mu}}$ where $\lambda = 0.1$ and $\mu = \frac{1}{\lambda} = 10$
  
  $P(X \ge 8) = e^{\frac{-8}{10}}$ 
```{r}
pexp(8, rate = 0.1, lower.tail = F)
```

$E(X) = \frac{1}{\lambda}$\
```{r}
1 / 0.1
```


$D(X)$

```{r}
sqrt(1/0.1^2)
```

 
### c.  What is the probability that the machine will fail after 8 years?.  Provide also the expected value and standard deviation.  Model as a binomial.  (Hint:  0 success in 8 years)   
  
  $b(n, p, k) =\binom{n}{k}p^kq^{n-k} $
  
  $P(K = 0) = b(8, 0.1, 0)$
 
```{r}
pbinom(0, 8, 0.1)

```
 
 $E(X) = np$
 
```{r}
n = 8
p = 0.1

n*p
```
 
 
$D(X) = \sqrt{npq}$
```{r}
n = 8
p = 0.1
q = 1- p
sqrt(n * p * q)
```


### d.  What is the probability that the machine will fail after 8 years?.  Provide also the expected value and standard deviation.  Model as a Poisson.   
  
 $P(X = k) \approx \frac{\lambda k}{k!}e^{−\lambda}$ where lambda $8 / 10 = 0.8$
 
 Failing after 8 years, that is 8 or more years -- $P(X \ge 8)$
 
 $\lambda$ is the expected number of successes in 8 years which is `0`.
 
```{r}
ppois(0, lambda=0.8, lower.tail = F)
```

$E(X) = \lambda = 0.8$

$D(X) = sqrt(\lambda)$
```{r}
sqrt(0.8)
```


Helpful Links:

[Problem Sets](https://math.dartmouth.edu/archive/m20f10/public_html/HW5Solutions.pdf)

[Poisson Distribution]([http://www.r-tutor.com/elementary-statistics/probability-distributions/poisson-distribution)
  
  