---
title: "DATA605 Wk9 Assignment 9"
author: "Javern Wilson"
date: "10/26/2019"
output: 
  html_document:
    theme: yeti
---

## 1. The price of one share of stock in the Pilsdorff Beer Company (see Exercise 8.2.12) is given by $Y_n$ on the *n*th day of the year. Finn observes that the diﬀerences $X_n = Y_n+1 −Y_n$ appear to be independent random variables with a common distribution having mean $\mu = 0$ and variance $\sigma^2= 1/4$. If $Y_1 = 100$, estimate the probability that $Y_{365}$ is 


Let $S^*_n = (S_n−nµ)/\sqrt{n}σ$ and 

$X_n = Y_{n+1} −Y_n$ therefore

$X_{364} = Y_{365}-Y_1$
 
$n = 364$


(a) $\ge100$. 
```{r}
1-pnorm((100-100)/(sqrt(364)*0.5))
```



(b) $\ge110$. 

```{r}
1-pnorm((110-100)/(sqrt(364)*0.5))
```



(c) $\ge120$. 

```{r}
1-pnorm((120-100)/(sqrt(364)*0.5))

```




## 2. Calculate the expected value and variance of the binomial distribution using the moment generating function. 


The Binomal distribution is given as:

$\binom{n}{x}p^x(1-p)^{n-x}$ or
$b(x; n, p) = \frac{n!}{x!(n − x)!}p^xq^{n−x}$ with $q = 1 − p$

Moment Generation function is given by 

\[ 
\begin{aligned}
M(t)&=\sum_{x = 0}^{n} e^{tx}\binom{n}{x}p^x(1-p)^{n-x}\\
&=\sum_{x = 0}^{n} e^{tx}\frac{n!}{x!(n − x)!}p^xq^{n−x}\\
&=\sum_{x = 0}^{n} (pe^t)^x\frac{n!}{x!(n − x)!}xq^{n−x}\\
&=(pe^t + 1 − p)^n\\
&=(pe^t + q)^n\\
\end{aligned}
\]


With the first moment we differentiate the MFG with respect to `t`


\[
\begin{aligned}
M'(t)&=n(pe^t + q)^{n-1}pe^t\\
&=npe^t(pe^t+q)^{n-1}
\end{aligned}
\]

Evaluating at t=0 we get the mean
\[\begin{aligned}M'(0)=E(X)&= np(q + p)^{n−1}\\
&=np(1-p+p)^{n-1}\\
&=np
\end{aligned}\]


For the second moment with t = 0

\[
\begin{aligned}
M''(t)= E(X^2)&=(n-1)npe^t(pe^t+q)^{n-2}pe^t + (npe^t)(pe^t + q)^{n-1}\\
&=np(n-1)(p+q)^{n-2}(p)+np(p+q)^{n-1}\\
&=np(n-1)p+np\\
&=(n^2p - np)p+np\\
&=n^2p^2-np^2 + np\\
&= np(np-p+1)\\
&=np(np+q)
\end{aligned}
\]

The Variance

\[
\begin{aligned}
V(X)&=E(X^2) - [E(X)]^2\\
&=np(np+q)- n^2p^2\\
&=n^2p^2+npq-n^2p^2\\
&=npq
\end{aligned}
\]


## 3. Calculate the expected value and variance of the exponential distribution using the moment generating function. 

The exponential function goes as follows:

$f_X^{(x)} = \lambda e^{-\lambda x}$

Using the MFG method

\[
\begin{aligned}M(t)&=
\left(\int_{-\infty}^{+\infty} e^{tx}f_X^{(x)} \; dx\right)\\
&=\left(\int_{0}^{\infty} e^{tx}\lambda e^{-\lambda x} \; dx\right)\\
&=\left(\int_{0}^{\infty} e^{x(t- \lambda)}\; dx\right)\\
&=\frac{\lambda}{t- \lambda}\bigg[e^{x(t- \lambda)}\bigg]_{0}^{\infty}\\
&=\frac{\lambda}{t- \lambda}\bigg[\displaystyle \lim_{x \to \infty} e^{x(t- \lambda)}- e^{(t- \lambda)0}\bigg]\\
&=\frac{\lambda}{t- \lambda}[0-1]\\
&=\frac{\lambda}{\lambda - t}\quad  where \quad t-\lambda < 0, t<\lambda\end{aligned} 

\]

*N.B* $M(t)$ can be written also as $\lambda(\lambda-t)^{-1}$

Finding the first moment

\[
\begin{aligned}
M'(0) = E(X)&= \lambda(\lambda-t)^{-1}\\
&=(-1)(\lambda)(\lambda-t)^{-2}(-1) \quad use \ of \ chain \ rule\\
&=\lambda(\lambda-t)^{-2}\\
&=\lambda(\lambda-0)^{-2}\\
&=\frac{\lambda}{(\lambda)^2}=\frac{1}{\lambda}
\end{aligned}
\]

Second derivative

\[
\begin{aligned}
M''(0)=E(X^2)&= \lambda(\lambda-t)^{2}\\
&=(-2)(\lambda)(\lambda-t)^{-3}(-1) \quad use \ of \ chain \ rule\\
&=2\lambda(\lambda-t)^{-3}\\
&=2\lambda(\lambda-0)^{-3}\\
&=\frac{2\lambda}{(\lambda)^3}= \frac{2\lambda}{\lambda^2}
\end{aligned}
\]

The Variance

\[
\begin{aligned}V(X) &= E(X^2) - [E(X)]^2\\
&=\frac{2}{\lambda^2} - \frac{1}{\lambda^2}\\
&=\frac{1}{\lambda^2}
\end{aligned}
\]

### Helpful Links

[Binomial](https://www.le.ac.uk/users/dsgp1/COURSES/MATHSTAT/5binomgf.pdf)

[Exponential (Video)](https://www.youtube.com/watch?v=QUAecbMv0y0)

[Exponential](https://bookdown.org/probability/beta/moment-generating-functions.html#moments-and-taylor-series)